\doctitle

### 2023-04-12 appendix/pinode_appendix.tex
Add PIKN chapter; Add todos across the doc

>>>>

\subsection{Burger's Equation: Compressibility of Linear Latent Space for PIKN}
\label{appendix:pikn_burgers}
\paragraph{Architecture} 
In the PIKN architecture for Burger's equation, the encoder is implemented as a a feed-forward network with the size of input layer set to be equal 128 (the number of spatial grid-points). It has two hidden layers with 512 neurons. The size of the output layer of the encoder -- the size of the latent dimension -- varies from 16 to 256 to study the latent space compressibility (see Figure~\ref{fig:pikn_burgers_compression}). The decoder's architecture mirrors the architecture of the encoder with the sizes of the inputs and outputs switched. An Adam optimizer has been applied for $500$ epochs for training. We use an adaptive learning rate: initially set to $10^{-4}$, it keeps decreasing by a factor of $0.5$ if no improvement has been made over the recent $20$ epochs, until it hits the minimal value of $10^{-7}$. \\
Each network was trained in two different ways. Namely, we set different values for the weights $\omega_1, \omega_2, \omega_3, \omega_4$ in the loss function, leading to two different training regimes:
\begin{itemize}
    \item $\omega_1=0, \omega_2=0, \omega_3=1, \omega_4=1$. These settings lead to a purely data-driven learning; it corresponds to green lines on Figure~\ref{fig:pikn_burgers_compression}.
    \item $\omega_1=1, \omega_2=1, \omega_3=1, \omega_4=1$. These settings represent a model that uses both data trajectories and collocations (a hybrid model); it corresponds to blue lines on Figure~\ref{fig:pikn_burgers_compression}.
\end{itemize}

\paragraph{Training} 
We sample $80000$ functions $\{u^{(1)}, u^{(2)}, \dots, u^{(80000)}\}$ to use them as collocations. Each function $u^{(j)}$ is a superposition of the Fourier modes that satisfy the periodic boundary conditions. We call them "harmonic" initial conditions, five examples of which are displayed on the middle-right pane of Figure~\ref{fig:pikn_burgers_compression}. We used $\sin{(kx)}$ and $\cos{(kx)}$ for $k=0, 1, 2, \dots$ evaluated on $x \in [-\pi; \pi]$-interval as basis functions. The value of $k$ is restricted to be no greater than $10$. To evaluate $u_t$ at each collocation point we evaluated the spatial derivatives $\{(u_{x}^{(i)}, u_{xx}^{(i)})\}_{i=1}^{80000}$ using a numerical spectral method. 
% \paragraph{Train Data}

For training trajectories, we use simulation data obtained from a solver base on spectral method. We run the simulation with $1024$ different initial states. Each initial state of $u$ is a Gaussian (bell-) curve with mean 0 and a randomly generated variance $\sigma \sim U(0.1,1)$, evaluated on $[-\pi; \pi]$-interval. Then we sample snapshots of the state $u$ with a temporal gap $\Delta t=0.1$ for $20$ steps (i.e. $p=20$). An example trajectory is displayed on the top-right pane of Figure~\ref{fig:pikn_burgers_compression}.

% \paragraph{Test Data} 
To evaluate and compare the performance of both models we use 200 trajectories: 100 trajectories with harmonic ICs and 100 trajectories with Gaussian ICs generated exactly in the same way as described above. Then we sample snapshots of the state $u$ with a temporal gap $\Delta t=0.1$ for $20$ steps (i.e. $p=20$).

\paragraph{Results} 
The results are aggregated in Table~\ref{table:pikn_burgers_performance} and visualized on Figure~\ref{table:pikn_burgers_performance}.

\begin{table}[t!]
    \centering
    \resizebox{\columnwidth}{!}{\input{tables/pikn_performance_table}}
    \caption{Performance measures for Burger's experiment depending on the dimensions of the latent space (Figure~\ref{fig:pikn_burgers_compression})}
    \label{table:pikn_burgers_performance}
\end{table}

### 2023-04-12 msr3.tex
Add PIKN chapter; Add todos across the doc

Linear Mixed-Effects (LME) models are a fundamental tool for modeling correlated data, 
including cohort studies, longitudinal data analysis, and meta-analysis.  Design and analysis of variable selection methods for LMEs is more difficult than for linear regression because LME models are nonlinear. In this work we 
propose a relaxation strategy and optimization methods that enable a wide range of variable selection methods for LMEs using both convex and nonconvex regularizers, including  $\ell_1$, Adaptive-$\ell_1$, SCAD, and $\ell_0$. 
The computational framework only requires the proximal operator for each regularizer to be available, and the implementation is available in an open source \texttt{python} package \texttt{pysr3}, consistent 
with the \texttt{sklearn} standard. 
The numerical results on simulated data sets indicate that the proposed strategy improves on the state of the art for both accuracy and compute time. 
The variable selection techniques are also validated on a real example using a data set on bullying victimization.

@@

\begin{table}[h]
\centering
\input{tables/table_glmm_current.tex}
\caption{\label{table:glmmlasso} Comparison of performance of MSR3-Fast for $\ell_1$ regularizer vs \texttt{glmmLasso}. MSR3-Fast executes 5 times faster in wall time and has higher accuracy of selecting correct covariates. 
%Importantly, the accuracy of \texttt{glmmLasso} is likely skewed downwards due to BIC selection criterion choosing dense ultimate models and due to the missing option to constrain the matrix $\Gamma$ to be diagonal. \texttt{lmmLasso} supports the diagonal specification of $\Gamma$ which translates into a competitive quality of selecting random effects. However, while finding sparse fixed effects, \texttt{lmmLasso} provides dense solutions for fixed effects $\beta$.
}
\end{table}

\subsection{Experiments on Real Data}
\label{sec:real}

@@

@@

\section{Software Implementation}
To ensure reproducibility of this research, all new algorithms have been implemented as a part of the \texttt{pysr3}\footnote{Available at \href{https://github.com/aksholokhov/pysr3}{https://github.com/aksholokhov/pysr3}} library. This library implements functionality for fitting linear mixed models and selecting covariates. The user interface was designed to be fully compliant with the standards\footnote{\href{https://scikit-learn.org/stable/developers/develop.html}{https://scikit-learn.org/stable/developers/develop.html}} of \texttt{sklearn} library to minimize learning time.

>>>>

\edit{Rewrite the MSR3 abstract into a summary of whole work with citations}
Linear Mixed-Effects (LME) models are a fundamental tool for modeling correlated data, 
including cohort studies, longitudinal data analysis, and meta-analysis.  Design and analysis of variable selection methods for LMEs is more difficult than for linear regression because LME models are nonlinear. In this work we 
propose a relaxation strategy and optimization methods that enable a wide range of variable selection methods for LMEs using both convex and nonconvex regularizers, including  $\ell_1$, Adaptive-$\ell_1$, SCAD, and $\ell_0$. 
The computational framework only requires the proximal operator for each regularizer to be available, and the implementation is available in an open source \texttt{python} package \texttt{pysr3}, consistent 
with the \texttt{sklearn} standard. 
The numerical results on simulated data sets indicate that the proposed strategy improves on the state of the art for both accuracy and compute time. 
The variable selection techniques are also validated on a real example using a data set on bullying victimization.

@@

\begin{table}[ht]
\centering
\resizebox{\columnwidth}{!}{\input{tables/table_glmm_current.tex}}

\caption{\label{table:glmmlasso} Comparison of performance of MSR3-Fast for $\ell_1$ regularizer vs \texttt{glmmLasso}. MSR3-Fast executes 5 times faster in wall time and has higher accuracy of selecting correct covariates. 
%Importantly, the accuracy of \texttt{glmmLasso} is likely skewed downwards due to BIC selection criterion choosing dense ultimate models and due to the missing option to constrain the matrix $\Gamma$ to be diagonal. \texttt{lmmLasso} supports the diagonal specification of $\Gamma$ which translates into a competitive quality of selecting random effects. However, while finding sparse fixed effects, \texttt{lmmLasso} provides dense solutions for fixed effects $\beta$.
}
\end{table}

\subsection{Sensitivity Analysis}
\label{ch:msr3_sensitivity_analysis}
\edit{Add sensitivity analysis from MSR3 practice paper}

\subsection{Application to Real Data: Anxiety and Depression as a Result of Bullying Victimization}
\label{sec:real}

@@

\subsection{Application to Real Data: COVID-19 Transmission Factor}
\edit{Add Covid example from my general}

@@

\section{Software Implementation}
\edit{Move pysr3 software paper here}
To ensure reproducibility of this research, all new algorithms have been implemented as a part of the \texttt{pysr3}\footnote{Available at \href{https://github.com/aksholokhov/pysr3}{https://github.com/aksholokhov/pysr3}} library. This library implements functionality for fitting linear mixed models and selecting covariates. The user interface was designed to be fully compliant with the standards\footnote{\href{https://scikit-learn.org/stable/developers/develop.html}{https://scikit-learn.org/stable/developers/develop.html}} of \texttt{sklearn} library to minimize learning time.

### 2023-04-12 pinode.tex
Add PIKN chapter; Add todos across the doc

\section{Introduction}

        Forecasting the behavior of a large-scale real-world system directly from first principles often requires solving highly-nonlinear governing equations such as high-dimensional ordinary differential equations (ODEs) or partial differential equations (PDEs). 
        High-fidelity simulations of such dynamical systems can become intractable especially if an online control algorithm requires multiple forecasts per second using a low-powered embedded device \cite{rowley2017model,lucia2004reduced,benner2015survey}. A situation like this arises, for example, when a smart heating, ventilation, and air conditioning (HVAC) system attempts to optimize the temperature distribution of the air in a room using only partial measurements \cite{farahmand2016learning,nabi2022robust}. At the moment of writing this paper such systems are incapable of real-time complex simulations, but they already can run low-dimensional pre-trained models, thus inviting the development of high-quality reduced order models (ROMs) \cite{otterness2017evaluation} . ROMs, therefore, are essential for enabling design optimization, uncertainty propagation, predictive models, and control for such dynamical systems \cite{brunton2022data,kutz2016dynamic,rowley2017model,jones2020characterising}

@@

        In this paper, we employ autoencoders to perform nonlinear model reduction along with NODE at the latent space to model complex and nonlinear dynamics. Our goal is to reduce their demand for data and improve overall forecasting stability under challenging training conditions. To achieve that we build on ideas from classical collocation methods of numerical analysis to embed knowledge from a known governing equation into the latent-space dynamics of a ROM, as described in Section 2. Then, in Section 3, we show that addition of our physics-informed loss allows for exceptional data supply strategies that improves the performance of ROMs in data-scarce settings, where training high-quality data-driven models is impossible. We demonstrate that such an approach not only reduces the need for large training data-sets and produces highly-accurate and long-term stable models, but also leads to the discovery of more compact latent spaces, which is especially important for applications in compressed sensing and control.

@@

\paragraph{Architecture} In this work we model $\psi$, $\psi^{-1}$, and $\bd{h}$ with fully-connected neural networks $\psi_\theta$, $\phi_\theta$, and $h_\theta$, respectively. Specifically, the pair ($\psi$, $\psi^{-1}$) is modelled with an auto-encoder $(\psi_\theta, \phi_\theta)$, and $\bd{h}$ is modelled with a fully-connected network $h_\theta$. Figure~\ref{fig:architecture} visualizes the architecture of the model.

@@

\paragraph{Collocations}
\label{sec:collocations_conditions} We define a collocation as pair $(\bd{\bar{x}},\, \bd{f}(\bd{\bar{x}}))$. Collocations are samples from the space $\XX \times Im_{f}(\XX)$, and they should satisfy three conditions, ordered by importance:
\begin{enumerate}
    \item \textbf{Simplicity}: $\bd{f}(\bar{\bd{x}}_j)$ should be computationally cheap to evaluate. It is especially important for PDE systems, where $\bd{f}$ may involve high-order derivatives.
    \item \textbf{Representativeness}: $\bar{\bd{x}}_j$ should cover the space of states where one aims to improve the model's performance or stability. Collocation points that a model might encounter and that are not represented by data snapshots are the best candidates.
    \item \textbf{Feasibility}: $\bar{\bd{x}}_j \in \XX$. In other words, $x_j$ should be an attainable state of the system. Collocation points outside of $\XX$ may downgrade the performance of the autoencoder by forcing it to be an invertible function on a domain outside of $\XX$ where a true mapping $\psi$ operates on.
\end{enumerate}
Thus, an optimal sampling procedure for collocations $\bd{\bar{x}}_j$ is domain-specific and should be designed given a particular system $\bd{f}$ and available data $\bd{x}_i$. We show examples of how these conditions can be implemented for real systems in the following sections.

@@

To the best of our knowledge, this is the first framework that combines non-linear latent-dynamics (Neural ODE), autoencoders, and a physics-informed loss term~(\ref{eq:loss_physics_informed}). Thus, we call our framework \textit{Physics-Informed Neural ODE}, or PINODE.

\section{Experiments}
\label{sec:exp}
The experiments section is organized as follows. First, to illustrate the ideas behind the framework we study its performance on a high-dimensional ODE -- a lifted Duffing oscillator. We show how a non-linear latent dynamics $\bd{h}(\bd{z})$ overcomes the limitations of DMD and Koopman networks from \cite{liu2022physics} by handling multiple basins of attraction within one model. We also show that using physics-informed loss is sufficient for reconstructing the behaviour for basins of attraction that are not represented by the data. Finally, we demonstrate that a purely data-driven model may be highly-accurate short-term and highly unstable long-term, even when the data is abundant, and show that the physics-informed approach improves long-term stability of such models by multiple orders of magnitude.

@@

\subsection{Burger's equation}
\label{sec:burger_eqn}
We now study the performance of our framework on Burger's equation with $[-\pi, \pi]$-periodic boundary conditions:
\begin{equation}
\begin{split}
\label{eq:burgers_equation}
    & u_t  + uu_x = \nu u_{xx} \\
    & u(-\pi, t) = u(\pi, t),\quad \forall t \in [0, T]
\end{split}
\end{equation}
where $u_t$, $u_x$, and $u_{xx}$ represent partial derivatives in time, the first, and second spatial derivatives, respectively. Burgers' equation is a PDE occurring in applications to acoustics, gas and fluid dynamics, and traffic flows \cite{burgers1948mathematical}. When $\nu$ is significantly smaller than one, the system exhibits strong non-linear behaviour and is called ``advection-dominated'', otherwise when $\nu$ is large the system is called ``diffusion-dominated''. 
In the case of the former, linear projection methods such as POD become inaccurate as the true solution space has a slow decaying Kolmogorov n-width, manifesting itself in slow decaying singular values \cite{peherstorfer2022breaking}. Therefore, in this section we focus on the advection-dominated Burger for which we set $\nu = 0.01$.

@@

\label{sec:compressibility}
\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/compressibility.pdf}
    \caption{PINODE Hybrid model utilized the latent space dimension 5 times more efficiently by MSE than PIKN Hybrid model when modelling low-viscosity (highly-nonlinear) Burger's equation (left frame). The difference in performance grows to x100, when forecasting two times farther than the train period (central frame). The reason behind PIKN's long-term instability is the precsence of eigenvalues with poitive real part in the latent dynamics matrix (right frame).}
    \label{fig:burgers_compressibility}
\end{figure}

@@

We compare the performance of three models: DMD, PIKN Hybrid, and PINODE Hybrid. First, we notice that DMD, despite achieving small loss values on training data ($\sim 10^{-3}$), does not perform well on test data. This observation is consistent with earlier works~(\cite{kalur2021robust,kutz2016dynamic}); it illustrates well that a combination of a linear encoder and a linear latent dynamics may not be sufficient for modelling highly-nonlinear phenonmena. Second, we notice that PINODE achieves better performance within a given latent space budget. For instance, for $m = 16$ (Figure~(\ref{fig:burgers_compressibility}), left pane), PINODE achieves $\sim 5$ times lower mean squared error than PIKN, which achieves the same performance only when $m = 512$. More importantly, PINODE maintains low error when predicting for a longer-term horizon (extrapolation in time), which is not the case for PIKN (Figure~(\ref{fig:burgers_compressibility}, right pane). This is a consequence of the latent-dynamics matrix ($h(z) = Lz$) of PIKN having eigenvalues with positive real parts, which implies long-term instability (Figure~(\ref{fig:burgers_compressibility}), right pane). Although there has been progress in the literature~\cite{kojimalearning}, further research is needed to understand (i) how to enforce stability constraints for PIKN, and (ii) why one does not need the same enforcement for PINODE to exhibit stable behaviour.

>>>>

\section{Introduction}
\edit{Write a big-picture intro to machine learning for physics}
\edit{Replace mentions of paper with work}
\subsection{Related Work}
\edit{Split original intro into prior works paragraphs: DL for Ph, ROMs, PIML}
\edit{Add physics-informed ML overview (PINN, DeepONet, NFO)}
        Forecasting the behavior of a large-scale real-world system directly from first principles often requires solving highly-nonlinear governing equations such as high-dimensional ordinary differential equations (ODEs) or partial differential equations (PDEs). 
        High-fidelity simulations of such dynamical systems can become intractable especially if an online control algorithm requires multiple forecasts per second using a low-powered embedded device \cite{rowley2017model,lucia2004reduced,benner2015survey}. A situation like this arises, for example, when a smart heating, ventilation, and air conditioning (HVAC) system attempts to optimize the temperature distribution of the air in a room using only partial measurements \cite{farahmand2016learning,nabi2022robust}. At the moment of writing this paper such systems are incapable of real-time complex simulations, but they already can run low-dimensional pre-trained models, thus inviting the development of high-quality reduced order models (ROMs) \cite{otterness2017evaluation} . ROMs, therefore, are essential for enabling design optimization, uncertainty propagation, predictive models, and control for such dynamical systems \cite{brunton2022data,kutz2016dynamic,rowley2017model,jones2020characterising}

@@

        In this paper, we employ autoencoders to perform nonlinear model reduction along with NODE at the latent space to model complex and nonlinear dynamics. Our goal is to reduce their demand for data and improve overall forecasting stability under challenging training conditions. To achieve that we build on ideas from classical collocation methods of numerical analysis to embed knowledge from a known governing equation into the latent-space dynamics of a ROM, as described in Section 2. Then, in Section 3, we show that addition of our physics-informed loss allows for exceptional data supply strategies that improves the performance of ROMs in data-scarce settings, where training high-quality data-driven models is impossible. We demonstrate that such an approach not only reduces the need for large training data-sets and produces highly-accurate and long-term stable models, but also leads to the discovery of more compact latent spaces, which is especially important for applications in compressed sensing and control.
        \edit{Add Koopman example to the intro}

        \edit{Add compressive sensing application to the intro}

@@

\paragraph{Architecture} In this work we model $\psi$, $\psi^{-1}$, and $\bd{h}$ with fully-connected neural networks $\psi_\theta$, $\phi_\theta$, and $h_\theta$, respectively. Specifically, the pair ($\psi$, $\psi^{-1}$) is modeled with an auto-encoder $(\psi_\theta, \phi_\theta)$, and $\bd{h}$ is modeled with a fully-connected network $h_\theta$. Figure~\ref{fig:architecture} visualizes the architecture of the model.

@@

\paragraph{Collocations}
\label{sec:collocations_conditions} We define a collocation as pair $(\bd{\bar{x}},\, \bd{f}(\bd{\bar{x}}))$. Collocations are samples from the space $\XX \times Im_{f}(\XX)$, and they should satisfy three conditions, ordered by importance:
\begin{enumerate}
    \item \textbf{Simplicity}: $\bd{f}(\bar{\bd{x}}_j)$ should be computationally cheap to evaluate. It is especially important for PDE systems, where $\bd{f}$ may involve high-order derivatives.
    \item \textbf{Representativeness}: $\bar{\bd{x}}_j$ should cover the space of states where one aims to improve the model's performance or stability. Collocation points that a model might encounter and that are not represented by data snapshots are the best candidates.
    \item \textbf{Feasibility}: $\bar{\bd{x}}_j \in \XX$. In other words, $x_j$ should be an attainable state of the system. Collocation points outside of $\XX$ may downgrade the performance of the auto-encoder by forcing it to be an invertible function on a domain outside of $\XX$ where a true mapping $\psi$ operates on.
\end{enumerate}
Thus, an optimal sampling procedure for collocations $\bd{\bar{x}}_j$ is domain-specific and should be designed given a particular system $\bd{f}$ and available data $\bd{x}_i$. We show examples of how these conditions can be implemented for real systems in the following sections.

@@

To the best of our knowledge, this is the first framework that combines non-linear latent-dynamics (Neural ODE), auto-encoders, and a physics-informed loss term~(\ref{eq:loss_physics_informed}). Thus, we call our framework \textit{Physics-Informed Neural ODE}, or PINODE.

\section{Experiments}
\label{sec:exp}
The experiments section is organized as follows. First, to illustrate the ideas behind the framework we study its performance on a high-dimensional ODE -- a lifted Duffing oscillator. We show how a non-linear latent dynamics $\bd{h}(\bd{z})$ overcomes the limitations of DMD and Koopman networks from \cite{liu2022physics} by handling multiple basins of attraction within one model. We also show that using physics-informed loss is sufficient for reconstructing the behavior for basins of attraction that are not represented by the data. Finally, we demonstrate that a purely data-driven model may be highly-accurate short-term and highly unstable long-term, even when the data is abundant, and show that the physics-informed approach improves long-term stability of such models by multiple orders of magnitude.

@@

\subsection{Burger's equation}
\label{sec:burger_eqn}
We now study the performance of our framework on Burger's equation with $[-\pi, \pi]$-periodic boundary conditions:
\begin{equation}
\begin{split}
\label{eq:burgers_equation}
    & u_t  + uu_x = \nu u_{xx} \\
    & u(-\pi, t) = u(\pi, t),\quad \forall t \in [0, T]
\end{split}
\end{equation}
where $u_t$, $u_x$, and $u_{xx}$ represent partial derivatives in time, the first, and second spatial derivatives, respectively. Burgers' equation is a PDE occurring in applications to acoustics, gas and fluid dynamics, and traffic flows \cite{burgers1948mathematical}. When $\nu$ is significantly smaller than one, the system exhibits strong non-linear behavior and is called ``advection-dominated'', otherwise when $\nu$ is large the system is called ``diffusion-dominated''. 
In the case of the former, linear projection methods such as POD become inaccurate as the true solution space has a slow decaying Kolmogorov n-width, manifesting itself in slow decaying singular values \cite{peherstorfer2022breaking}. Therefore, in this section we focus on the advection-dominated Burger for which we set $\nu = 0.01$.

@@

\label{sec:compressibility}
\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/compressibility.pdf}
    \caption{PINODE Hybrid model utilized the latent space dimension 5 times more efficiently by MSE than PIKN Hybrid model when modeling low-viscosity (highly-nonlinear) Burger's equation (left frame). The difference in performance grows to x100, when forecasting two times farther than the train period (central frame). The reason behind PIKN's long-term instability is the presence of eigenvalues with positive real part in the latent dynamics matrix (right frame).}
    \label{fig:burgers_compressibility}
\end{figure}

@@

We compare the performance of three models: DMD, PIKN Hybrid, and PINODE Hybrid. First, we notice that DMD, despite achieving small loss values on training data ($\sim 10^{-3}$), does not perform well on test data. This observation is consistent with earlier works~(\cite{kalur2021robust,kutz2016dynamic}); it illustrates well that a combination of a linear encoder and a linear latent dynamics may not be sufficient for modeling highly-nonlinear phenomena. Second, we notice that PINODE achieves better performance within a given latent space budget. For instance, for $m = 16$ (Figure~(\ref{fig:burgers_compressibility}), left pane), PINODE achieves $\sim 5$ times lower mean squared error than PIKN, which achieves the same performance only when $m = 512$. More importantly, PINODE maintains low error when predicting for a longer-term horizon (extrapolation in time), which is not the case for PIKN (Figure~(\ref{fig:burgers_compressibility}, right pane). This is a consequence of the latent-dynamics matrix ($h(z) = Lz$) of PIKN having eigenvalues with positive real parts, which implies long-term instability (Figure~(\ref{fig:burgers_compressibility}), right pane). Although there has been progress in the literature~\cite{kojimalearning}, further research is needed to understand (i) how to enforce stability constraints for PIKN, and (ii) why one does not need the same enforcement for PINODE to exhibit stable behavior.

    \begin{figure}[t!]
        \centering
        \includegraphics[width=\textwidth]{figures/burgers_compression.png}
        \caption{The right plots give examples of data snapshots used: trajectories with bell-curve ICs (top) and harmonic ICs (bottom). The middle-right pane shows harmonic collocations used by hybrid models in addition to the snapshots. The left plot compares the prediction errors (MSE) of two models, data-driven and hybrid with 128-dimensional latent spaces, on harmonic initial conditions that were not present in the trained data. The shaded regions represent 95\% confidence intervals based on 100 test trajectories. The middle plot shows the prediction error (MSE) on each type of test data for a variety of models with different latent-space sizes. We see that the use of harmonic collocations significantly improves the model performance on unseen harmonic ICs without increasing the errors on bell-curve ICs.}
        \label{fig:pikn_burgers_compression}
    \end{figure}

    However, we observe that PIKN benefits from using physics-informed loss as well. In particular, we show that one can use collocations to improve model's performance on types of initial conditions that are missing in the available training data. To illustrate that we train two models. The data-driven model only uses 1024 trajectories with bell-curve initial conditions (ICs) (we provide an example at the top-right frame of the Figure~\ref{fig:pikn_burgers_compression}). The hybrid model additionally observes 80000 harmonic collocations formed by summing first 10 sinusoidal modes with random coefficients (Figure~\ref{fig:pikn_burgers_compression}, middle-right frame), for which we evaluate $u_t$ analytically using Equation~\ref{eq:burgers_equation}. Next, we evaluate the performance of both models using unseen trajectories with both harmonic and bell-curve ICs, 100 trajectories each. We observe that the Hybrid model predicts the sinusoidal trajectories ~10 times better than the data-driven one (Figure~\ref{fig:pikn_burgers_compression}, left frame, shown for the 128-dimensional latent-space model). Since neither models had any trajectories of that type in its training set we conclude that the difference in performance comes from using harmonic collocations. We also note that better performance of the hybrid model on harmonic ICs does not come at an expense of worse performance on bell-curve ICs, as shown in the central frame of Figure~\ref{fig:pikn_burgers_compression}. This evidence suggests that one can improve a model's extrapolation power  by supplementing its training with sufficiently diverse set of collocations, especially when additional simulations are expensive to obtain but the collocations are cheap to generate. The details of the network's architecture and training procedure are provided in the Appendix \ref{appendix:pikn_burgers}.
